Should I Airbnb My Property?

### 2017-03-21 abb_Writeup.Rmd
GA edits, intro litreview

---
title: "Should I Airbnb My Property?"
author: "Andy Krause and Gideon Aschwanden"
date: "March 10, 2017"
output:
  pdf_document: default
  html_document: default
header-includes: 
   - \usepackage{setspace}\doublespacing
   - \usepackage{caption}
   - \captionsetup[table]{skip=10pt}
geometry: margin=3.3cm
bibliography: abb.bib
---

@@

With the rise of Airbnb and other self-hosting accommodation platforms^[Flipkey.com, VRBO.com, Homeaway.com, Tripping.com, Stayz.com.au as examples.], it is easier than ever for property owners and investors to consider short-term rental (STR) as a potential leasing strategy for a revenue generating residential property. Should they though? Or rather, is it more profitable to do so when compared to a more traditional long-term rental (LTR)? And further, are certain types of properties in specific locations more likely to show a preference for a short-term leasing strategy over long-term one.

To address these questions and to provide some guidance to potential and current investors and owners, we examine the short- and long-term residential rental market in Melbourne, Australia. Melbourne has a thriving short-term rental market with over 20,000 total properties listed on the Airbnb portal, the source of our short-term market data. We adopt a scenario-based approach that compares observed annual revenues for short-term rental (STR) properties with an estimate of their expected long-term rental revenues. Our results show that the feasibility of a greater return through a short-term rental -- what we have termed ``Short-Term Preference'' -- is dependent on property type, product, location and host policies.  In short, the answer to whether or not a owner or investor should opt for a short-term leasing strategy: it depends.

@@

Existing research into short-term rental markets is sparse. However, with the arrival of Airbnb, and the controversies surrounding it, the topic is creating interest both in real estate and economics as well as in a wide variety of fields as evidenced by a growth in recent publications and on-going working papers.  Within the extant work, the focus remains centered on rural and/or amenity-based STRs such as beach-front and ski-resort access properties.  This focus resulted from the fact that, until a few years ago, the STR market itself operated almost wholly within this sphere.  The shift to small, urban dwellings within the STR market is relatively new and explains the dearth of past work in this direction.

@@

Though issues of second homes have been primarily treated as a rural phenomenom for decades  [@gallent2007second], a wave of ``new urban tourism'' Fuller and Michel [@fuller2014stop] has re-focused some of the problems of short-term rentals on large, urban environments. With this move to urban areas, issues such as nuisance, property rights and regulation have naturally arisen.  As a result, the short-term rental market has garnered significant interest in the public policy and legal sphere as of late. As many second homes are not permitted for or regulated as commercial entities, STRs can have serious effects on neighbors in the way of externalies such as noise and congestion [@frost2006commentary].  While strict regulation is often proposed, doing so may constititute regulatory takings [@jefferson2015can].  Other, rights-based, regulatory approaches have been suggested [@miller2014transferable].  Many of the key legal decisions regarding short-term rentals, particulary though Airbnb, are still in progress and much remains to be settled in this area in terms of the regulatory and legal environment in which STRs are allowed to operate.

@@

Australia implements a particular tax policy, often termed `negative gearing', that allows for favourable tax treatment to real estate investors who experience greater costs than revenues on investment properties.^[See @fane2004negative for a complete explanation of negative gearing.] To date, this policy only applies to long-term rental properties.  While this policy can influence the net revenue (post tax) of a property, the negative gearing policy requires knowledge of the investor's personal income situation.  As we do not have access to this information in the marketplace, we will not be considering any effects of this policy in our analysis here.

@@

Leasing costs $L$ are the fees paid to the rental agency to find and manage tenants over the life of the long-term lease.  In Melbourne, rates generally range from 5\% to 7\% per annum.  We have assumed leasing costs of 6\% of the total gross revenue of the long-term lease contract.

>>>>

---
title: "Should I Airbnb My Property?"
author: "Andy Krause and Gideon Aschwanden"
date: "March 17, 2017"
output:
  pdf_document: default
  html_document: default
header-includes: 
   - \usepackage{setspace}\doublespacing
   - \usepackage{caption}
   - \captionsetup[table]{skip=10pt}
geometry: margin=3.3cm
bibliography: abb.bib
---

@@

With the rise of Airbnb and other self-hosting accommodation platforms^[Flipkey.com, VRBO.com, Homeaway.com, Tripping.com, Stayz.com.au as examples.], it is easier than ever for property owners and investors to consider short-term rental (STR) as a potential leasing strategy for a revenue generating residential property. Should they though? Or rather, is it more profitable to do so when compared to long-term rental (LTR)? And further, are certain types of properties in specific locations more likely to show a preference for a short-term leasing strategy over long-term one.

To address these questions and to provide some guidance to potential and current investors and owners, we examine the short- and long-term residential rental market in Melbourne, Australia. Melbourne has a thriving short-term rental market with over 20,000 total properties listed on the Airbnb portal, the source of our short-term market data.

We adopt a scenario-based approach that compares observed annual revenues for short-term rental (STR) properties with an estimate of their expected long-term rental revenues. Our results show that the feasibility of a greater return through a short-term rental -- what we have termed ``Short-Term Preference'' -- is dependent on property type, product, location and host policies.  In short, the answer to whether or not a owner or investor should opt for a short-term leasing strategy: it depends.

@@

Existing research into short-term rental markets is sparse. However, with the arrival of Airbnb, and the controversies surrounding it, the topic is creating interest both in real estate and economics as well as in a wide variety of fields as evidenced by a growth in recent publications and on-going working papers.  Within the existing work, the focus remains centered on rural and/or amenity-based STRs such as beach-front and ski-resort access properties.  This focus resulted from the fact that, until a few years ago, the STR market itself operated almost wholly within these market segments. The shift to small, urban dwellings within the STR market is relatively new and explains the dearth of past work in this direction.

@@

Though issues of second homes have been primarily treated as a rural phenomenom for decades  [@gallent2007second], a wave of ``new urban tourism'' Fuller and Michel [@fuller2014stop] has re-focused some of the problems of short-term rentals on large, urban environments. With this move to urban areas, issues such as nuisance, property rights and regulation have naturally arisen.  As a result, the short-term rental market has garnered significant interest in the public policy and legal sphere as of late. As many second homes are not permitted for or regulated as commercial entities, STRs can have serious effects on immediate neighbors in the way of externalies such as noise and congestion [@frost2006commentary].  While strict regulation is often proposed, doing so may constititute regulatory takings [@jefferson2015can].  Other, rights-based, regulatory approaches have been suggested [@miller2014transferable].  Many of the key legal decisions regarding short-term rentals, particulary though Airbnb, are still in progress and much remains to be settled in this area in terms of the regulatory and legal environment in which STRs are allowed to operate.

@@

Australia implements a particular tax policy, often termed `negative gearing', that allows for income tax deduction to real estate investors who experience greater costs than revenues on investment properties.^[See @fane2004negative for a complete explanation of negative gearing.] To date, this policy only applies to long-term rental properties.  While this policy can influence the net revenue (post tax) of a property, the negative gearing policy requires knowledge of the investor's personal income situation.  As we do not have access to this information in the marketplace, we will not be considering any effects of this policy in our analysis here.

@@

Leasing costs $L$ are the fees paid to the rental agency to find and manage tenants over the life of the long-term lease.  In Melbourne, agency rates generally range from 5\% to 7\% per annum.  We have assumed leasing costs of 6\% of the total gross revenue of the long-term lease contract.
